Requantization of time-dependent mean field for pairing collective motion in superfluid nuclei

### 2019-01-01 thesis/thesis.tex
thesis

In early days, the low-lying excited $0^+$ states in most of nuclei had been considered as $\beta$ vibration or two-phonon excitation associated with quadrupole vibrations. However, plenty of recent experimental data indicate that not only quadrupole correlations but also pairing correlations play important roles in such excited $0^+$ states. 
%From theoretical viewpoint, 
%excited $0^+$ states in superfluid nuclei correspond to not only quadrupole vibrational mode, the elementary mode of excitation of quadrupole collective motions, but also pairing vibrational mode, the elementary mode of excitation of pairing collective motions. 
The competition between quadrupole correlations and pairing correlations causes complicate and mysterious structure for the low-lying excited $0^+$ states. 
We aim to understand such excited $0^+$ states based on time-dependent mean-field (TDMF) theory. Because quantum fluctuation is large in low-lying excitation region, we need to consider the collective motion in large-amplitude regime. However, the theoretical framework to describe pairing collective motions in large-amplitude regime is not well developed. The nature of pairing collective motions in low-lying excitation region is still unclear.

We consider the theoretical framework to describe large-amplitude collective motions associated with nuclear pairing. In general, the low-lying excited states can be obtained in two steps: (1) solve TDMF equation to obtain the classical trajectories; (2) quantize such classical trajectories to obtain the excited states. The step 2 is called ``requantization''. Because the theoretical framework for the requantization is not well developed, we focus on the step 2 and discuss it in detail. 
%To simplify the system, 
Apart from realistic nuclear system, 
we consider the pairing collective motions in multi-level pairing model. We divide the pairing system into two classes, integrable system and non-integrable system. The former one corresponds to two-level system and the later one corresponds to the system more than three-level. Because the description of requantization is different between the two, we consider the requantization methods for each case, respectively.

In integrable system, the requantization of TDMF is particular feasible because we can easily find the periodic orbits. We apply three different requantization methods, canonical quantization, Fourier decomposition of the time-dependent observables, and stationary phase approximation (SPA) to the path integral. Comparing each requantization method with exact solution in various two-level system, respectively, SPA turns out to be most accurate between the three and can describe the excited states quantitatively, especially for the description of two-particle transition strength. Because the requantized wave functions from SPA are given in terms of the microscopic degrees of freedom, the transition matrix elements can be calculated directly.

Actually, SPA cannot be applied to non-integrable system straightforwardly because finding the periodic orbits is extremely difficult task. We propose a new theoretical approach which enables SPA to apply to non-integrable system, that is, combining SPA with adiabatic self-consistent collective coordinate (ASCC) method (ASCC+SPA). The concept is that we employ ASCC to extract an integrable subspace from original TDMF phase space, and apply SPA to obtain the excited states. 
%The detailed derivation is shown in the text.
After the application to the low-lying excited $0^+$ states in three-level system and Pb isotopes system, we find that ASCC+SPA can properly describe the pairing collective motions.

We also examine the performance of the pairing collective model which assumes pairing gap parameter and gauge angle as collective coordinates. We check the one-to-one correspondence between the collective coordinates from TDMF dynamics itself and those in the collective model. We find the collective coordinate describing pairing vibration has no one-to-one correspondence between the two, in both integrable and non-integrable system. The applicability of the collective model is limited for the nuclear pairing.

We summarize the achievements in this thesis. We find a reasonable requantization method, SPA, from the study in integrable system. Using SPA, we construct a theoretical framework of ASCC+SPA to requantize the pairing collective motions in non-integrable system. Also, we demonstrate that it is necessary to describe the pairing vibration in the collective coordinate extracted from TDMF degrees of freedom itself, instead of the phenomenological collective coordinate, pairing gap parameter. With the development of computational power and numerical techniques, we think 
the ASCC+SPA is the promising tool to elucidate the structure of mysterious low-lying excited $0^+$ states in nuclei, in the near future.

@@

We apply the SPA approach to the two-level pairing model.
The invariant measure in SU(2)$\otimes$SU(2) is 
\begin{align}
d\mu(Z) &= \prod_{\alpha=1,2} \frac{\Omega_{\alpha}+1}{\pi}(1+|Z_{\alpha}|^2)^{-2}d({\rm Re}Z_{\alpha})d({\rm Im}Z_{\alpha}) \nonumber \\
  &= \prod_{\alpha=1,2} \frac{\Omega_{\alpha}+1}{4\pi}d(\cos{\theta_{\alpha}})d\chi_{\alpha} \nonumber \\
	&= \left(\prod_{\alpha=1,2} \frac{1+\Omega_{\alpha}^{-1}}{2\pi}\right)
             d\Phi dJ d\phi dj .
\end{align}
In the last equation, we transform the canonical coordinates
by Eqs. (\ref{phi}) and (\ref{pi}).
Since the particle number $J=N/2$ and the total energy $E$ are invariant,
the two-level pairing model is integrable.
Thus, we can construct the semiclassical wave function using
Eq. (\ref{semi_wave_func}).
From (\ref{Lagrangian2}), the action integral is given by
\begin{align}
	\mathcal{T}_k(\Phi,\phi;J)
	&= J\Phi + \int^t_0
	j(t') \frac{d\phi}{dt'} dt' \nonumber \\
	&\equiv \mathcal{T}_{N,E_k}(\Phi,t) ,
\end{align}
where the integration is performed on the $k$-th
closed trajectory of Eq. (\ref{EBK}), and
the variables $(\phi,j)$ are transformed into $(t,E)$.
%where $\pi(t)$ can be solved from the right side of (\ref{TDHFB_equation}). (i.e. 2D Runge-Kutta method) \par
The semiclassical wave function fulfilling the EBK quantization condition
becomes
\begin{align}
	\ket{\psi_k^J} &\propto \oint d\Phi \oint dt
	e^{i\mathcal{T}_{J,E_k}(\Phi, t)}
	\ket{\Phi,t}_{N,E_k} \label{semi_wave_func0} \\
  &\propto \sum_{m=0}^{J}C_m^{(E_k,J)}\ket{S_1,-S_1+m;S_2,-S_2+(J-m)} ,
	\label{semi_wave_func1}
\end{align}
with $S_{\alpha}=(\Omega_{\alpha}-\nu_{\alpha})/2$ and the coefficients
\begin{align}
  C_m^{(E_k,J)} =& \binom{J}{m}\int_0^T dt \nonumber \\
 &\times \exp{\left( i\int^t j(t') \dot{\phi}(t') dt'-i(J/2-m)\phi(t) \right)} \nonumber \\ 
 &\times A(j_1,S_1,m)A(j_2,S_2,J-m) ,\\
  A(j,S,m) &= \left(\frac{j}{2S}\right)^{m/2}\left(\frac{2S-j}{2S}\right)^{S-m/2}\sqrt{\frac{(2S)!m!}{(2S-m)!}}, \hspace{5mm} \nonumber
\end{align}
where $T$ is the period of the closed trajectory.
The TDHFB-requantized wave functions (\ref{semi_wave_func1}) are
eigenstates of the total particle number.
This is due to the integration over the global gauge angle $\Phi$,
which makes the particle number projection not only for the ground state
but also for excited states.
See Appendix \ref{wave_func1} for detailed derivation of Eq.~(\ref{semi_wave_func1}).

@@

\subsection{Canonical quantization}
\label{sec:canonical}
In the pairing collective model, the canonical quantization with Pauli's prescription
was adopted in previous studies \cite{BBPK70,GPBW85, ZPPRS99, P07}.
Assuming magnitude and phase of the pairing gap as collective coordinates,
a collective Hamiltonian was constructed in the second order in momenta.
Then, the Hamiltonian was quantized by the canonical quantization
with Pauli's prescription.
In this section, we apply similar quantization method to
the TDHFB Hamiltonian (\ref{Hamiltonian2}).
The main difference is that the collective canonical variables are
not assumed in the present case, but are obtained from the TDHFB dynamics
itself.\par
It is not straightforward to apply Pauli's prescription
to the present case, 
because the TDHFB Hamiltonian (\ref{Hamiltonian2})
is not limited to the second order in momenta.
%We need to keep the Hamiltonian Hermitian after 
In the present study, 
we adopt a simple symmetrized ordering, as
\begin{align}
	\mathcal{H}(\hat{\phi},\hat{j};\hat{J})
	=& \sum_{\alpha=1,2} \epsilon_{\alpha}\nu_{\alpha} + \sum_{\alpha=1,2} 2\epsilon_{\alpha}j_{\alpha} - g\sum_{\alpha=1,2} \left( (\Omega_{\alpha}-\nu_{\alpha}) j_{\alpha} - j_{\alpha}^2 +\frac{j_{\alpha}^2}{\Omega_{\alpha}-\nu_{\alpha}} \right) \nonumber \\
	&- g \left\{ \sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})}\cos{\hat{\phi}} \right. \nonumber \\
	&\left. + \cos{\hat{\phi}}\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} \right\} .
\label{canonical_quantized_H}
\end{align}
To fulfill the canonical quantization condition
\begin{align}
 [\hat{\Phi},\hat{J}]=i, \quad\quad [\hat{\phi},\hat{j}]=i ,
\end{align}
as in (\ref{j_l}), $j_{\alpha}$ contain $J$ and $j$ which
are replaced by
\begin{equation}
	\hat{J} = -i\frac{\partial}{\partial\Phi},\quad\quad
	\hat{j} = -i\frac{\partial}{\partial\phi} .
\end{equation}
Since $\Phi$ is a cyclic variable, 
we write the collective wave function $\Psi(\Phi,\phi)$ as eigenstates
of the number of $J^{\pi}=0^+$ pairs $J$ in a separable form
\begin{equation}
  \Psi_k^{(J)}(\Phi,\phi) = 
	\frac{1}{\sqrt{2\pi}}e^{iJ\Phi}\psi_k^{(J)}(\phi) .
\end{equation}
Then, the problem is reduced to the one-dimensional Schr\"{o}dinger
equation for the motion in the relative angle $\phi$.
%pairing vibrational mode can be obtained by solving
the Schr\"{o}dinger equation is
\begin{equation}
	H\left( \phi,-i\frac{d}{d\phi};J \right)
	\psi_k^{(J)}(\phi) = E_k^{(J)}\psi_k^{(J)}(\phi) .
	\label{Schroedinger_eq}
\end{equation}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \caption{
The same as Fig.~\ref{fig:N50Pad} but for $N=98\rightarrow 100$.}
 \label{fig:N100Pad}
\end{figure}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
	\caption{The same as Fig.~\ref{fig:N50Pad} but for $N=6\rightarrow 8$
	with $\Omega=8$.
}
 \label{fig:N8Pad}
\end{figure}

\begin{figure}[htb]
 \begin{center}
  \includegraphics[height=0.6\textwidth,,angle=-90]{images/N16ex_energy_wo_adiabatic.eps}
 \end{center}
	\caption{The same as Fig.~\ref{fig:N8energy} but for
	$N=2\Omega=16$.
}
 \label{fig:N16energy}
\end{figure}

\begin{figure}[thb]
 \begin{minipage}{1\hsize}
 \begin{center}
   \includegraphics[height=0.45\textwidth,angle=-90]{images/N16Xeq0p5occ_wo_adiabatic.eps}
   \includegraphics[height=0.45\textwidth,angle=-90]{images/N16Xeq2occ_wo_adiabatic.eps}
 \end{center}
 \end{minipage}
	\caption{The same as Fig.~\ref{fig:N8_occ} but for
	$N=2\Omega=16$.
}
 \label{fig:N16_occ}
\end{figure}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
	\caption{The same as Fig.~\ref{fig:N50Pad} but for $N=14\rightarrow 16$
	with $\Omega=8$.
}
 \label{fig:N16Pad}
\end{figure}

@@

After deriving the collective subspace $\Sigma_2$, 
we perform the quantization according to the SPA in Sec.~\ref{sec:SPA}.
Calculating a trajectory in the $(q^1,p_1)$ space,
we can identify a series of states $\{\ket{\Phi,J;q^1(t),p_1(t)}\}$
on the trajectory,
in the form of Eq. (\ref{coherent})
with parameters $Z_\alpha$ given at $(\Phi,J,q^1,p_1)$ and $q^a=p_a=0$ for
$a\geq 3$.
%Due to $J$ is conserved quantity, the action integral can be divided into two terms
Since the variables $(\Phi,J)$ and $(q^1,p_1)$ are separable,
we may take closed trajectories independently in $(\Phi,J)$ and $(q^1,p_1)$
sectors, which we denote here as $C_\Phi$ and $C_1$, respectively.
The action integral is given by
\begin{align}
\mathcal{T}(\Phi,J;q^1,p_1)
=& \int_{C_\Phi} \braket{\Phi(t),J;q^1,p_1|i\frac{\partial}{\partial t}|\Phi(t),J;q^1,p_1} dt \nonumber \\
&+ \int_{C_1} \braket{\Phi,J;q^1(t),p_1(t)|i\frac{\partial}{\partial t}|\Phi,J;q^1(t),p_1(t)} dt
 \nonumber \\
	=& J\Phi + \int_{C_1} \sum_\alpha j^\alpha d\chi_\alpha
	\nonumber \\
	\equiv& \mathcal{T}_\Phi(J,\Phi) +\mathcal{T}_1(q^1,p_1;J) .
%	\label{tau}
\end{align}
In fact, the gauge-angle dependence is formally given as
\begin{equation}
  \ket{\Phi,J;q^1,p_1} = e^{-i\Phi \hat{J}} \ket{J;q^1,p_1},
\end{equation}
where 
\begin{align}
	\hat{J} = \frac{\hat{N}-\nu}{2} .
\end{align}
Then, the action for the trajectory $C_1$ can be also expressed as
$\mathcal{T}(q^1,p_1;J)=
 \int_{C_1} \braket{J;q^1,p_1|i\frac{\partial}{\partial t}|J;q^1,p_1} dt$.
%\begin{align}
%  \mathcal{T}(\Phi,J;q,p) &= \mathcal{T}_{\rm intr}(t) + J\Phi, 
%\end{align}
%%where the intrinsic action integral $\mathcal{T}_{\rm intr}(t)$ becomes
%The important point is $\mathcal{T}_{\rm intr}(t) \ne \int p dq$ at each $t$. For closed TDHFB trajectory, only $\mathcal{T}_{\rm intr}(t)$ contributes to the excited states.

@@

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=60mm,angle=-90]{images/intra_trans.eps}
 \end{center}
 \begin{center}
  \includegraphics[width=60mm,angle=-90]{images/inter_trans.eps}
 \end{center}
\caption{Calculated strength of pair-addition transition (\ref{BPad})
from $N=14$ to $N=22$. 
The solid (dashed) lines correspond to the ASCC+SPA (exact) calculation.
The horizontal line indicates the particle number of the initial states.
The upper panel shows the intraband transitions,
$0_1^+\to 0_1^+$ and $0_2^+\to 0_2^+$,
while the lower panel shows the interband transitions,
$0_1^+ \to 0_2^+$ and $0_2^+\to 0_1^+$.
}
 \label{3levelPad}
\end{figure}

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Properties of superfluid ground state in pairing model}
The superfluid ground state correspond to the local minimum point in energy surface. The time-dependent coherent state described in (\ref{coherent}) attributes to BCS trail wave function in $\chi_{\alpha}=0$. For the simplicity, here, we only consider the even-particle system with $\nu=0$ (without Pauli blocking).

>>>>

The low-lying excited $0^+$ states in most of nuclei had been traditionally considered as either $\beta$ vibration or two-phonon excitation associated with quadrupole vibrations. However, plenty of recent experimental data indicate that not only quadrupole correlations but also pairing correlations play important roles in such excited $0^+$ states. 
%From theoretical viewpoint, 
%excited $0^+$ states in superfluid nuclei correspond to not only quadrupole vibrational mode, the elementary mode of excitation of quadrupole collective motions, but also pairing vibrational mode, the elementary mode of excitation of pairing collective motions. 
The competition between quadrupole correlations and pairing correlations causes complex and mysterious structure for the low-lying excited $0^+$ states. 
We aim to understand such excited $0^+$ states based on the time-dependent mean-field (TDMF) theory. Because quantum fluctuation is large in low-lying excitation region, we need to treat the collective motion in large-amplitude regime. However, the theoretical framework to describe pairing collective motions in the large-amplitude regime is not yet established. The nature of low-lying pairing collective motions is still unclear.

We aim at developing a new theoretical framework to describe large-amplitude collective motions associated with nuclear pairing. In our approach, the low-lying excited states are obtained in two steps: (1) solve TDMF equation to obtain the classical trajectories; (2) quantize such classical trajectories to obtain the excited states. The step 2 is called ``requantization''. 
First, we consider the pairing collective motions in multi-level pairing model. The pairing models can be classified into two classes, integrable and non-integrable systems. The former one corresponds to the two-level system and the later one corresponds to the system with three levels or more.

In the integrable system, the requantization of TDMF is particularly feasible because we can easily find the periodic orbits. We apply three different requantization methods, canonical quantization, Fourier decomposition of the time-dependent observables, and stationary phase approximation (SPA) to the path integral. Comparing each requantization method with the exact solution, respectively, SPA turns out to be the most accurate among the three and can describe the excited states quantitatively, including two-particle transition strength. Because the requantized wave functions from SPA are given in terms of the microscopic degrees of freedom, the transition matrix elements can be calculated directly.

Unfortunately, the application of the SPA to the non-integrable system is not straightforward because finding the periodic orbits is an extremely difficult task. We propose a new theoretical approach which makes the SPA to be applicable to non-integrable systems, that is, combining the SPA with the adiabatic self-consistent collective coordinate (ASCC) method (ASCC+SPA). The concept is that we employ the ASCC to extract an integrable subspace from original TDMF phase space, and apply the SPA to the integrable subspace. 
%The detailed derivation is shown in the text.
We apply the ASCC+SPA to the low-lying excited $0^+$ states in three-level system and Pb isotopes system. We find that the ASCC+SPA can properly describe the pairing collective motions.

We also examine the validity of the pairing collective model which assumes the pairing gap parameter and gauge angle as the phenomenological collective coordinates. We check the one-to-one correspondence between the collective coordinates self-consistently determined with the ASCC and those assumed in the collective model. We find the collective coordinate describing pairing vibration has no one-to-one correspondence between the two, in both integrable and non-integrable systems. The applicability of the collective model is limited for the nuclear pairing.

%We summarize the achievements in this thesis. We find a reasonable requantization method, SPA, from the study in integrable system. Using SPA, we construct a theoretical framework of ASCC+SPA to requantize the pairing collective motions in non-integrable system. Also, we demonstrate that it is necessary to describe the pairing vibration in the collective coordinate extracted from TDMF degrees of freedom itself, instead of the phenomenological collective coordinate, pairing gap parameter. 
With the development of computational power and numerical techniques,
the ASCC+SPA is a promising tool to elucidate the structure of mysterious low-lying excited $0^+$ states in nuclei, in near future.

@@

We apply the SPA approach to the two-level pairing model.
The invariant measure in SU(2)$\otimes$SU(2) is 
\begin{align}
d\mu(Z) &= \prod_{\alpha=1,2} \frac{2S_{\alpha}+1}{\pi}(1+|Z_{\alpha}|^2)^{-2}d({\rm Re}Z_{\alpha})d({\rm Im}Z_{\alpha}) \nonumber \\
  &= \prod_{\alpha=1,2} \frac{2S_{\alpha}+1}{4\pi}d(\cos{\theta_{\alpha}})d\chi_{\alpha} \nonumber \\
	&= \left(\prod_{\alpha=1,2} \frac{1+(2S_{\alpha})^{-1}}{2\pi}\right)
             d\Phi dJ d\phi dj .
\end{align}
In the last equation, we transform the canonical coordinates
by Eqs. (\ref{phi}) and (\ref{pi}).
Since the particle number $J=(N-\nu)/2$ and the total energy $E$ are invariant,
the two-level pairing model is integrable.
Thus, we can construct the semiclassical wave function using
Eq. (\ref{semi_wave_func}).
From (\ref{Lagrangian2}), the action integral is given by
\begin{align}
	\mathcal{T}_k(\Phi,\phi;J)
	&= J\Phi + \int^t_0
	j(t') \frac{d\phi}{dt'} dt' \nonumber \\
	&\equiv \mathcal{T}_{J,E_k}(\Phi,t) ,
\end{align}
where the integration is performed on the $k$-th
closed trajectory of Eq. (\ref{EBK}), and
the variables $(\phi,j)$ are transformed into $(t,E)$.
%where $\pi(t)$ can be solved from the right side of (\ref{TDHFB_equation}). (i.e. 2D Runge-Kutta method) \par
The semiclassical wave function fulfilling the EBK quantization condition
becomes
\begin{align}
	\ket{\psi_k^J} &\propto \oint d\Phi \oint dt
	e^{i\mathcal{T}_{J,E_k}(\Phi, t)}
	\ket{\Phi,t}_{J,E_k} \label{semi_wave_func0} \\
  &\propto \sum_{m=0}^{J}C_m^{(E_k,J)}\ket{S_1,-S_1+m;S_2,-S_2+(J-m)} ,
	\label{semi_wave_func1}
\end{align}
with $S_{\alpha}=(\Omega_{\alpha}-\nu_{\alpha})/2$ and the coefficients
\begin{align}
  C_m^{(E_k,J)} =& \binom{J}{m}\int_0^T dt \nonumber \\
 &\times \exp{\left( i\int^t j(t') \dot{\phi}(t') dt'-i(J/2-m)\phi(t) \right)} \nonumber \\ 
 &\times A(j_1,S_1,m)A(j_2,S_2,J-m) ,\\
  A(j,S,m) &= \left(\frac{j}{2S}\right)^{m/2}\left(\frac{2S-j}{2S}\right)^{S-m/2}\sqrt{\frac{(2S)!m!}{(2S-m)!}}, \hspace{5mm} \nonumber
\end{align}
where $T$ is the period of the closed trajectory.
The TDHFB-requantized wave functions (\ref{semi_wave_func1}) are
eigenstates of the total particle number.
This is due to the integration over the global gauge angle $\Phi$,
which makes the particle number projection not only for the ground state
but also for excited states.
See Appendix \ref{wave_func1} for detailed derivation of Eq.~(\ref{semi_wave_func1}).

@@

\subsection{Canonical quantization}
\label{sec:canonical}
In the pairing collective model, the canonical quantization with Pauli's prescription
was adopted in previous studies \cite{BBPK70,GPBW85, ZPPRS99, P07}.
Assuming magnitude and phase of the pairing gap as collective coordinates,
a collective Hamiltonian was constructed in the second order in momenta.
Then, the Hamiltonian was quantized by the canonical quantization
with Pauli's prescription.
In this section, we apply similar quantization method to
the TDHFB Hamiltonian (\ref{Hamiltonian2}).
The main difference is that the collective canonical variables are
not assumed in the present case, but are obtained from the TDHFB dynamics
itself.\par
It is not straightforward to apply Pauli's prescription
to the present case, 
because the TDHFB Hamiltonian (\ref{Hamiltonian2})
is not limited to the second order in momenta.
The application of Pauli's prescription for the adiabatic TDHFB Hamiltonian (\ref{ATDHFB_Hamiltonian}) is discussed in Appendix \ref{can-trans}.
Here, 
we adopt a simple symmetrized ordering, as
\begin{align}
	\mathcal{H}(\hat{\phi},\hat{j};\hat{J})
	=& \sum_{\alpha=1,2} \epsilon_{\alpha}\nu_{\alpha} + \sum_{\alpha=1,2} 2\epsilon_{\alpha}j_{\alpha} - g\sum_{\alpha=1,2} \left( (\Omega_{\alpha}-\nu_{\alpha}) j_{\alpha} - j_{\alpha}^2 +\frac{j_{\alpha}^2}{\Omega_{\alpha}-\nu_{\alpha}} \right) \nonumber \\
	&- g \left\{ \sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})}\cos{\hat{\phi}} \right. \nonumber \\
	&\left. + \cos{\hat{\phi}}\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} \right\} .
\label{canonical_quantized_H}
\end{align}
To fulfill the canonical quantization condition
\begin{align}
 [\hat{\Phi},\hat{J}]=i, \quad\quad [\hat{\phi},\hat{j}]=i ,
\end{align}
as in (\ref{j_l}), $j_{\alpha}$ contain $J$ and $j$ which
are replaced by
\begin{equation}
	\hat{J} = -i\frac{\partial}{\partial\Phi},\quad\quad
	\hat{j} = -i\frac{\partial}{\partial\phi} .
\end{equation}
Since $\Phi$ is a cyclic variable, 
we write the collective wave function $\Psi(\Phi,\phi)$ as eigenstates
of the number of $J^{\pi}=0^+$ pairs $J$ in a separable form
\begin{equation}
  \Psi_k^{(J)}(\Phi,\phi) = 
	\frac{1}{\sqrt{2\pi}}e^{iJ\Phi}\psi_k^{(J)}(\phi) .
\end{equation}
Then, the problem is reduced to the one-dimensional Schr\"{o}dinger
equation for the motion in the relative angle $\phi$.
%pairing vibrational mode can be obtained by solving
the Schr\"{o}dinger equation is
\begin{equation}
	H\left( \phi,-i\frac{d}{d\phi};J \right)
	\psi_k^{(J)}(\phi) = E_k^{(J)}\psi_k^{(J)}(\phi) .
	\label{Schroedinger_eq}
\end{equation}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N100Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \caption{
The same as described in the caption of Fig.~\ref{fig:N50Pad} but for $N=98\rightarrow 100$.}
 \label{fig:N100Pad}
\end{figure}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
  \includegraphics[width=70mm,angle=-90]{images/N8Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
	\caption{The same as described in the caption of Fig.~\ref{fig:N50Pad} but for $N=6\rightarrow 8$
	with $\Omega=8$.
}
 \label{fig:N8Pad}
\end{figure}

\begin{figure}[htb]
 \begin{center}
  \includegraphics[height=0.6\textwidth,,angle=-90]{images/N16ex_energy_wo_adiabatic.eps}
 \end{center}
	\caption{The same as described in the caption of Fig.~\ref{fig:N8energy} but for
	$N=2\Omega=16$.
}
 \label{fig:N16energy}
\end{figure}

\begin{figure}[thb]
 \begin{minipage}{1\hsize}
 \begin{center}
   \includegraphics[height=0.45\textwidth,angle=-90]{images/N16Xeq0p5occ_wo_adiabatic.eps}
   \includegraphics[height=0.45\textwidth,angle=-90]{images/N16Xeq2occ_wo_adiabatic.eps}
 \end{center}
 \end{minipage}
	\caption{The same as described in the caption of Fig.~\ref{fig:N8_occ} but for
	$N=2\Omega=16$.
}
 \label{fig:N16_occ}
\end{figure}

@@

\begin{figure}[t]
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_CQ.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_FD.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.3\hsize}
 \begin{center}
 \includegraphics[width=70mm,angle=-90]{images/N16Pad_SPA.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
	\caption{The same as described in the caption of Fig.~\ref{fig:N50Pad} but for $N=14\rightarrow 16$
	with $\Omega=8$.
}
 \label{fig:N16Pad}
\end{figure}

@@

After deriving the collective subspace $\Sigma_2$, 
we perform the quantization according to the SPA in Sec.~\ref{sec:SPA}.
Calculating a trajectory in the $(q^1,p_1)$ space,
we can identify a series of states $\{\ket{\Phi,J;q^1(t),p_1(t)}\}$
on the trajectory,
in the form of Eq. (\ref{coherent})
with parameters $Z_\alpha$ given at $(\Phi,J,q^1,p_1)$ and $q^a=p_a=0$ for
$a\geq 3$.
%Due to $J$ is conserved quantity, the action integral can be divided into two terms
Since the variables $(\Phi,J)$ and $(q^1,p_1)$ are separable,
we may take closed trajectories independently in $(\Phi,J)$ and $(q^1,p_1)$
sectors, which we denote here as $C_\Phi$ and $C_1$, respectively.
The action integral is given by
\begin{align}
\mathcal{T}(\Phi,J;q^1,p_1)
=& \int_{C_\Phi} \braket{\Phi(t),J;q^1,p_1|i\frac{\partial}{\partial t}|\Phi(t),J;q^1,p_1} dt \nonumber \\
&+ \int_{C_1} \braket{\Phi,J;q^1(t),p_1(t)|i\frac{\partial}{\partial t}|\Phi,J;q^1(t),p_1(t)} dt
 \nonumber \\
	=& J\Phi + \int_{C_1} \sum_\alpha j^\alpha d\chi_\alpha
	\nonumber \\
	\equiv& \mathcal{T}_\Phi(J,\Phi) +\mathcal{T}_1(q^1,p_1;J) .
%	\label{tau}
\end{align}
In fact, the gauge-angle dependence is formally given as
\begin{equation}
  \ket{\Phi,J;q^1,p_1} = e^{-i\Phi \hat{J}} \ket{J;q^1,p_1},
\end{equation}
where 
\begin{align}
	\hat{J} = \frac{\hat{N}-\nu}{2} = \sum_{\alpha} (\hat{S}_{\alpha}^0 + S_{\alpha}).
\end{align}
Then, the action for the trajectory $C_1$ can be also expressed as
\begin{align}
\mathcal{T}(q^1,p_1;J)=
 \int_{C_1} \braket{J;q^1,p_1|i\frac{\partial}{\partial t}|J;q^1,p_1} dt.
\end{align}
%\begin{align}
%  \mathcal{T}(\Phi,J;q,p) &= \mathcal{T}_{\rm intr}(t) + J\Phi, 
%\end{align}
%%where the intrinsic action integral $\mathcal{T}_{\rm intr}(t)$ becomes
%The important point is $\mathcal{T}_{\rm intr}(t) \ne \int p dq$ at each $t$. For closed TDHFB trajectory, only $\mathcal{T}_{\rm intr}(t)$ contributes to the excited states.

@@

\begin{figure}[htbp]
 \begin{minipage}{1\hsize}
 \begin{center}
  \includegraphics[height=0.45\textwidth,angle=-90]{images/intra_trans.eps}
  \includegraphics[height=0.45\textwidth,angle=-90]{images/inter_trans.eps}
 \end{center}
 \end{minipage}
\caption{Calculated strength of pair-addition transition (\ref{BPad})
from $N=14$ to $N=22$. 
The solid (dashed) lines correspond to the ASCC+SPA (exact) calculation.
The horizontal line indicates the particle number of the initial states.
The left panel shows the intraband transitions,
$0_1^+\to 0_1^+$ and $0_2^+\to 0_2^+$,
while the right panel shows the interband transitions,
$0_1^+ \to 0_2^+$ and $0_2^+\to 0_1^+$.
}
 \label{3levelPad}
\end{figure}

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Extra discussion in two-level pairing model}

\section{Phase transition point}
\label{two-level}
We consider the phase transition point between the superfluid state and the normal state. The superfluid ground state corresponds to the local minimum point in potential energy surface. The time-dependent coherent state (\ref{coherent}) attributes to BCS trail wave function in $\chi_{\alpha}=0$. For the sake of simplicity, here, we only consider the even-particle system with $\nu=0$ (without Pauli blocking).
